---
title: "Project Models"
format: html
editor: visual
---

This file is for building and testing machine learning models.

```{r setup}

#Load tools for data cleaning
library(tidyverse)

#Load and preview dataset for analysis
model_data <- read.csv("Transportation Data.csv")
#view(model_data)

## Use model_data for machine learning models below
```

## Variables + Data Info

```{r}
names(model_data)
dim(model_data)
```

## Model 1 - Linear Regression

### Model 1 - 1

Outcome variable - sustainable_modes_total, Predictor - median age.

For every 1 unit increase in the median age, there is a -.21 decrease of people taking a sustainable mode of transit within a census tract (holding all other variables constant).

```{r}
#all modes combined (pub transit, carpool, walking, biking)
lm.fit <- lm(sustainable_modes_total ~ age_medianE, data = model_data)
summary(lm.fit)
#public transit
lm.fit2 <- lm(transport_mode_public_transitE ~ age_medianE, data = model_data)
summary(lm.fit2)
#carpool
lm.fit3 <- lm(transport_mode_carpoolE ~ age_medianE, data = model_data)
summary(lm.fit3)
#walk
lm.fit4 <- lm(transport_mode_walkE ~ age_medianE, data = model_data)
summary(lm.fit4)
#bike
lm.fit5 <- lm(transport_mode_bikeE ~ age_medianE, data = model_data)
summary(lm.fit5)
```

```{r}
lm_am<- lm(sustainable_modes_total ~ median_monthly_housing_costsE + race_whiteE,
 data = model_data)
summary(lm_am)
```

### Model 1 - 2

Outcome - sustainable_modes_total, Predictor - State.

States with higher prob of taking sustainable modes of transit are Alaska,California, Colorado, Connecticut, DC, Hawaii, Illinois, Maine, Maryland, Massachusetts, Minnesota, Montana, Nevada, New Jersey, New York, Oregon, Pennsylvania, Rhode Island, Vermont, Virginia, Washington, Wyoming.

\*Note - with coefficients greater than 2. Also cool to see that West Virginia, Tennesee, Mississippi, and Arkansas have the smallest coefficients ( \< 1)

```{r}
lm.fit_1 <- lm(sustainable_modes_total ~ State, data = model_data)
summary(lm.fit_1)
```

### Model 1 - 3

For every 1 unit increase in \# of transit stops, we see an .11 increase in the amount of folks taking alternative transit.

```{r}
lm.fit_2 <- lm(sustainable_modes_total ~ Number.of.Transit.Stops, data = model_data)
summary(lm.fit_2)
```

### Model 1 - 4

For every 1 unit increase in \# of intersections, we see a -0.003151 decrease in amount of folks taking sustainable modes of transit.

```{r}
lm.fit_3 <- lm(sustainable_modes_total ~ Number.of.Intersections, data = model_data)
summary(lm.fit_3)
```

### Model 1 - 5

With every 1 unit increase in the Intersections per acre, we see an 11.5 unit increase in amount of people taking sustainable modes of transit.

```{r}
lm.fit_4 <- lm(sustainable_modes_total ~ Intersection.Per.Acre, data = model_data)
summary(lm.fit_4)
```

### Model 1 - 6

For every 1 unit increase in \# stops per acre, we see a 69.3 increase in amount of folks taking sustainable modes of transit.

```{r}
lm.fit_5 <- lm(sustainable_modes_total ~ Stops.Per.Acre, data = model_data)
summary(lm.fit_5)
```

## Model 2 - Multiple Linear Regressions

### Model 2 - 1

Intersections per acre and stops per acre are noteworthy here.

```{r}
mlr.fit <- lm(sustainable_modes_total ~ Number.of.Intersections + Number.of.Transit.Stops + Intersection.Per.Acre + Stops.Per.Acre, data = model_data)
summary(mlr.fit)
plot(mlr.fit)
```

### Model 2 -2

Outcomes - (1) sustainable_modes_total, (2) bike, (3) public transit

Predictors - racial identity

```{r}
mlr.fit2 <- lm(sustainable_modes_total ~ race_whiteE + race_blackE + race_native_americanE + race_pacific_islanderE + race_otherE + race_two_or_moreE + race_hispanicE, data = model_data)
summary(mlr.fit2)
```

```{r}
mlr.fit2_bike <- lm(transport_mode_bikeE ~ race_whiteE + race_blackE + race_native_americanE + race_pacific_islanderE + race_otherE + race_two_or_moreE + race_hispanicE, data = model_data)
summary(mlr.fit2_bike)
```

```{r}
mlr.fit2_pub_transit <- lm(transport_mode_public_transitE ~ race_whiteE + race_blackE + race_native_americanE + race_pacific_islanderE + race_otherE + race_two_or_moreE + race_hispanicE, data = model_data)
summary(mlr.fit2_pub_transit)
```

### Model 2 - 3

This fit takes really long - many counties.

```{r}
mlr.fit_many <- lm(sustainable_modes_total ~ race_whiteE + race_blackE + race_native_americanE + race_pacific_islanderE + race_otherE + race_two_or_moreE + race_hispanicE + pop_density + median_household_incomeE + median_monthly_housing_costsE + Number.of.Intersections + Number.of.Transit.Stops,
 data = model_data)
summary(mlr.fit_many)
```

## Model 3 - Linear Discriminant Analysis

### Model 3 - 1

```{r}

```

## Model X

```{r}

```

## Model X - Unsupervised Clustering

We are looking to potentially find clusters of census tracts to recommend policies to, in an effort to increase their sustainable mode of transportation usage rates. Clusters that are similar to each other in terms of demographics or infrastructure, but have higher rates of non-automobile usage may offer policies that can be cross-implemented into these census tracts and find success.

```{r}
num.only_data <- na.omit(model_data[,c(2,11,13,16,19,20,21,23,28,29)]) ### Need to get rid of any rows with na's, removed 1607 rows for upcoming PCA analysis. ##race_whiteE, race_native_american_E, race_other_E,age_medianE,median_household_incomeE, median_monthly_housing_costsE, ints.per.acre, transit.stops.per.acre
#model.data.matrix <- data.matrix(scale(num.only_data)) 

### PCA needs to be scaled with mean of 0, SD of 1. scale() does exactly that. prcomp() also needs a matrix to run, data.matrix() does this.

#matrix.dist <- as.dist(1 - cor(t(model.data.matrix))) 
### This gives us a correlation matrix that can be read as a distance matrix for prcomp() to chew through.

###ERROR - correlate matrix was too large, will demonstrate concept with sample of 1000 census tracts.
set.seed(10)
samp.vct <- sample(nrow(num.only_data), 1000)

sample.data <- num.only_data[samp.vct,c(8,9)]
samp.data.matrix <- data.matrix(scale(sample.data))

samp.dist <- as.dist(1 - cor(t(samp.data.matrix)))

samp.hier <- hclust(samp.dist, method = "complete")
cut.hier <- cutree(samp.hier, k = 5)


plot(samp.hier, main = "Complete Linkage with Correlation-Based Distance", xlab = "", sub = "")
plot(cut.hier, main = "Complete Linkage with Correlation-Based Distance, K = 8", xlab = "", sub = "")

cluster.df <- data.frame(cmdscale(samp.dist), factor(cut.hier))

names(cluster.df) <- c("x", "y", "cluster")
cluster.df$model <- rownames(cluster.df)


ggplot(cluster.df, aes(x, y)) + geom_point(aes(colour = cluster), size=3)
```

## Model X - Support Vector Machines

```{r}
#load package for support vector machines
library(e1071)

#Generate a histogram of sustainable travel mode usage to select the threshold for classification.
hist(model_data$sustainable_modes_total)

#Adjusting the data for a classification method by turning sustainable_modes_total (percentage) into lables Low (less than 10%) and High (10% or more). Using 10% since the majority are less than that, but a small portion are above.
model_data_svm <- mutate(
  model_data,
  sustainable_modes_class = case_when(
    sustainable_modes_total < 10 ~ "low",
    sustainable_modes_total >= 10 ~ "high"
  )
)

svm.fit()
  

```
